### 2011-04-12 Text/paper.tex
Wrote the numerical benchmark section. Need to obtain more results. Probably for the aspect ratio as well.

\section{Numerical benchmark approach}
\label{sec:numerical:benchmark}
The main discussion here follows and is based on the two-dimensional benchmark approach
\cite{kuzmin-binary2d}. It was indicated that the benchmark layout should have certain
dimensions to conduct simulations. The classical Bretherton benchmark layout is represented in Fig.
\ref{fig:classical:benchmark}. It describes the gas finger propagation through the liquid media.
The film thickness in this case is measured at the inlet. In lattice Boltzmann framework such
formulation has certain challenges \cite{kuzmin-binary2d}. Some of them are attributed to the
dynamic coupling of the inlet and outlet conditions \cite{giavedoni-numerical}. Instead, we
proposed the numerical benchmark indicated in Fig. \ref{fig:lbm:benchmark}. The dimensions of the
channel are chosen as $H_{\mathrm{eff}}\times H_{\mathrm{eff}} \times 15 H_{\mathrm{eff}}$. The
initialized bubble length is taken as $5 H_{\mathrm{eff}}$.\citet{giavedoni-numerical} showed that
the film stabilizes at the distances of $2.6-4.0$ diameters
from the front tip depending on the Reynolds number.  We chose to measure the film thickness in the
middle of the bubble, which is located at least at a distance of $2.5-3$ channel heights from the 
bubble tip. Moreover, the film thickness examination, see Section
\ref{section:film:thickness:variation}, shows that for the capillary range of interest, i.e.
$0.05\leq Ca \leq 1$, the standard deviation of the film thickness from the bubble middle film
thickness is around $7$
percent for small capillary numbers ($Ca=0.05$) and less than $1$ percent for larger ones ($Ca
\geq 0.1$). %at the distances one channel height from the rear tip upto one channel

This length of the bubble was shown as
sufficient for the film thickness to stabilize.

\begin{figure}[ht]
\includegraphics*[bb=153 610 405 717,width=\textwidth]{Figures/benchmark_classical.eps} 
\caption{The classical benchmark. The semiinfinite gas bubble
propagates through the liquid media. \label{fig:classical:benchmark}}
\end{figure}[ht]
\begin{figure}
\includegraphics*[bb=152 480 410 713,width=\textwidth]{Figures/benchmark_lbm.eps}
\caption{The classical benchmark. The semiinfinite gas bubble
propagates through the liquid media. \label{fig:lbm:benchmark}}
\end{figure}

 In order to avoid the mutual

It was indicated that the film thickness should be at least as twice large
as the interface thickness. The interface thickness in the continuous formulation of the
free-energy binary liquid formulation occupies $3-5$ nodes. Therefore, the film should be
resolved at least as $6-10$ nodes.

As it was indicated before \cite{heil-threedim}, the diagonal bubble radius is different from the
axial bubble radius for $Ca\leq \widehat{Ca}\approx 0.05$. The radii coincide for larger values of
$Ca$, but the value of bubble radii is $R_{diag}=R_{axis}=0.99 a$, where $a$ is the side length of
the square microchannel. Therefore, taking the minimal requirement for the film thickness to be
resolved, i.e. $6$ lattice nodes, one can obtain the grid size as $600\times 600 \times $ to
simulate

Non axisymmetric case is attributed to the case where the axial radius is different from the
diagonal radius and the bubble has the non-circular shape in the channel crossection. 
{\color{red}
\citet{cerro-space} indicate that because of asymmetry the flow in the corner can predominate the
flow in the side of the capillary thus changing the qualitative differencies depending on the
gravity direction. ``Because of the mixing, the gas-liquid mass tranfer coefficient are 2-5 times
larger fro upflow than for downflow''. 
}

Most multiphase lattice Boltzmann models \cite{swift, Shan-chen:extended} resolve
the interface using continuous interface methods where the interface spans over several grid nodes.
Such a representation
brings issues of the film thickness resolution versus interface
resolution -- the diffuse interface should be dealt with in such a way as to have a
negligible effect on the physics of the film. In our previous work \citet{kuzmin-binary2d} the grid
resolution was examined. It was found that the film thickness should be at least as twice as the
interface for the simulation to be accurate.

Therefore, the major governing parameter for microchannel Bretherton flows is not the density ratio,
but the
viscosity ratio. Thus, the applicability of the binary liquid model can be examined for the
simulation of the square shaped capillaries.

 Moreover, the results based on uniform density, as in the case of the LBM
binary liquid model, are in good agreement with other simulations
\cite{giavedoni-numerical,heil-bretherton} in which inertia effects were taken into account. The
same indicated in \cite{gupta-review} the  classical Bretherton problem is stated for
gas and liquid, which are of
significantly
different densities and viscosities.

Resolving the interface on a fine level for capillary numbers smaller than
$0.003$ is computationally expensive even in the two-dimensional case
(the amount of memory necessary to perform the simulation would be of the order
of tens or hundreds of gigabytes).  This work is therefore based on a
comparison of the results obtained with the known data
validated for the Bretherton bubble flow.  In this paper, we present
techniques for initialization of the simulations and discuss optimal
parameter ranges for the binary liquid LB model, such as the binary liquid parameters, the viscosity
ratio and the grid resolution. The emphasis is to obtain correct flow physics, especially in terms
of film thicknesses.

The VOF technique is the continuous
interface method based on a non-uniform grid. It is of particular interest for us to compare it
with the continuous interface method by the LBM implementation.

Thus, there is the effect of
inertia on the bubble formation.

Recent experimental results by \citet{shikazono-square} for square shaped capillaries indicate the
following dependency on the capillary number $Ca$ and the Weber number $We$:
\begin{equation}
\begin{aligned}
&R_{diag}=1.171-\frac{2.43 Ca^{2/3}}{1+7.28 Ca^{2/3}-0.255 We^{0.215}}\\
&R_{axis}=
\begin{cases}
1, &R_{diag}>1\\
R_{diag}, &R_{diag}\leq 1.
\end{cases}
\end{aligned}
\end{equation}
In the present
simulations the largest Reynolds number we achieve is less than $1$ and the capillary number is of
order of unity. Thus, it allows to neglect 
inertia effects.

\section{Steady-state case}
We performed different simulations to understand a number of steps required for the problem to come
on the steady-state. The grid to be simulated is $52\mathrm{x}52\mathrm{x}750$ which simulates the
quarter of the channel with the initial
width as $6.5$ lattice Boltzmann units together with $\frac{\mathrm{d}P}{\mathrm{d}x}=1.6
\mathrm{x}10^{-6}$. The simulation was performed for $140000$,$150000$,$160000$, $170000$ and
$180000$ iterations. The results are summarised in Table \ref{table:steady:state}. Whole the
capillary number varies only in the 4th digit, radiuses are one percent accurate.

>>>>

\section{Numerical benchmark approach}
\label{sec:numerical:benchmark}
The main discussion here follows and is based on the two-dimensional benchmark approach
\cite{kuzmin-binary2d}. It was indicated that the benchmark layout should have certain
dimensions to conduct simulations. The classical Bretherton benchmark layout is represented in Fig.
\ref{fig:classical:benchmark}. It describes the gas finger propagation through the liquid media.
The film thickness in this case is measured at the inlet. In lattice Boltzmann framework such
formulation has certain challenges \cite{kuzmin-binary2d}. Some of them are attributed to the
dynamic coupling of the inlet and outlet conditions \cite{giavedoni-numerical}. Instead, we
proposed the numerical benchmark indicated in Fig. \ref{fig:lbm:benchmark}. The dimensions of the
channel are chosen as $H_{\mathrm{eff}}\times H_{\mathrm{eff}} \times 15 H_{\mathrm{eff}}$. The
initialized bubble length is taken as $5 H_{\mathrm{eff}}$. \citet{giavedoni-numerical} showed that
the film stabilizes at the distances of $2.6-4.0$ diameters
from the front tip depending on the Reynolds number. \citet{heil-threedim} measure the bubble radii at the distance as $5.5 H_{\mathrm{eff}}$ to be sufficient for $Ca\leq 10$. In this work the conducted simulations capillary number is not larger than unity $Ca\leq 1$.  Thus, the film thickness is chosen to be measured in the
middle of the bubble, which is located at least at a distance of $2.5-3$ channel heights from the 
bubble tip.%at the distances one channel height from the rear tip upto one channel

For simplicity, the periodic boundary conditions are applied. Thus, the bubble train is simulated. To avoid the mutual influence of bubbles on each other the channel length was chosen as $15 H_{\mathrm{eff}}$. Periodic boundary conditions imply as well that the fluid is accelerated by a body force in the framework of the LBM but not with the pressure difference. Therefore, we can not address simulations with upward and downward flows \citet{cerro-bubble}. It is assumed that the gravity effects are negligible and the flow is in the horizontal direction driven by the body force.  
\begin{figure}[ht]
\includegraphics*[bb=153 610 405 717,width=\textwidth]{Figures/benchmark_classical.eps} 
\caption{The classical benchmark. The semiinfinite gas bubble
propagates through the liquid media. \label{fig:classical:benchmark}}
\end{figure}[ht]
\begin{figure}
\includegraphics*[bb=152 480 410 713,width=\textwidth]{Figures/benchmark_lbm.eps}
\caption{The classical benchmark. The semiinfinite gas bubble
propagates through the liquid media. \label{fig:lbm:benchmark}}
\end{figure}

It was indicated \cite{kuzmin-binary2d} that the film thickness for two-dimensional simulations should be at least as twice large
as the interface thickness. The interface thickness in the continuous formulation of the
free-energy binary liquid formulation occupies $3-5$ nodes. Therefore, the film should be
resolved at least as $6-10$ nodes.

\cite{heil-threedim} specify that the diagonal bubble radius is different from the
axial bubble radius for $Ca\leq \widehat{Ca}\approx 0.05$. The radii coincide for larger values of
$Ca$, but the value of bubble radii is $R_{diag}=R_{axis}=0.99 H_{\mathrm{eff}}$. Therefore, taking the minimal requirement for the film thickness to be
resolved, i.e. $6$ lattice nodes, one can obtain the grid size as $600\times 600 \times 9000 $ to properly resolve the binary liquid microchannels bubbles motion. This size implies highly computational resource to be available for simulations.

One can argue that the non-uniform grid can be used. However, the deposited film thickness is the function of the curvatures of the front bubble tip \cite{bretherton}. Therefore, the grid needs to be refined just at the interface between gas and liquid. It is complicated to achieve for the dynamic systems where a bubble moves in the streamwise direction.

Also, the binary liquid model is used. The model has an uniform density. A part of discussion is addressed in the introduction that to fully address the Bretherton problem one needs to take into the account gravitational, viscous, inertial, and surface tension forces. However,  \citet{shikazono-square} indicated that for square shaped capillaries radii have the
following dependency on the capillary number $Ca$ and the Weber number $We$:
\begin{equation}
\label{radii:experimental}
\begin{aligned}
&R_{diag}=1.171-\frac{2.43 Ca^{2/3}}{1+7.28 Ca^{2/3}-0.255 We^{0.215}}\\
&R_{axis}=
\begin{cases}
1, &R_{diag}>1\\
R_{diag}, &R_{diag}\leq 1.
\end{cases}
\end{aligned}
\end{equation}
Equation (\ref{radii:experimental}) shows that a change in the Reynolds number $0\leq Re \leq 100$ implies the change of the radii from $R_{diag}=R_{axis}=0.96$ to $0.93$ for $Ca=0.1$ and from $R_{diag}=R_{axis}=0.88$ to $0.85$ for $Ca=0.1$. The error of $3$ percent of the channel height is acceptable for current simulations where the Reynolds number is less than {\color{red}$100$}.

Thus, having limitations of the computational memory and performance requirements the focus of the present work is a feasibility study of the binary liquid lattice Boltzmann method for the microchannel simulations with square crossections conducted in the moderate capillary number range $0.1 \leq Ca \leq 1.0$. We base our work on comparison with the already available studies.

To further reduce the computational overhead only quarter of the channel is simulated, see Appendix \ref{appendix}. The next section further examines the steady-state approach, the dependency of the radii on the moderate capillary number values $0.1\leq Ca \leq 1.0$, the velocity pattern, the relative bubble velocity and the aspect ratio.

\section{Results}

\section{Steady-state approach}
We performed different simulations to understand a number of steps required for the problem to come
on the steady-state. The grid to be simulated is $52\mathrm{x}52\mathrm{x}750$ which simulates the
quarter of the channel with the initial
width as $6.5$ lattice Boltzmann units together with $\frac{\mathrm{d}P}{\mathrm{d}x}=1.6
\mathrm{x}10^{-6}$. The simulation was performed for $140000$,$150000$,$160000$, $170000$ and
$180000$ iterations. The results are summarised in Table \ref{table:steady:state}. Whole the
capillary number varies only in the 4th digit, radiuses are one percent accurate.
